
 p1 <- mean(v)
 #one sample formula
p1
p0 <- 0.8
n1 <- 30
z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
p <- pnorm(z, lower.tail = TRUE)
p

#two sample formula
pstar <- (p1+p2)/2

z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
z

p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
p

## I cold logically move the 2-tailed version up to become the first level, and then use if to negate that if n2 and p2 were null
```{r}
Z.prop.test <- function(p1, n1, p0, p2=NULL, n2=NULL, alternative="two.sided", conf.level=0.95){
n1xp1.assumption.test <- n1*p1
n1x1minusp1.assumption.test<- n1*(1-p1)

if (n1xp1.assumption.test<5 |n1x1minusp1.assumption.test*(1-p1)<5) {
      warning("WARNING: not valid to assume normal distribution. (n1*p1<5 or n1*(1-p1)<5) ")
      
}


      
#if (p2== NULL | n2==NULL){
      
      z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
            z
            
      if (alternative == "less"){
           
            p <- pnorm(z, lower.tail = TRUE)
            p
      }
      if (alternative == "greater"){
            
            p <- pnorm(z, lower.tail = FALSE)
            p
      }
      if (alternative == "two.tailed"){
           
            p <- 2*pnorm(z)
            p
      }
#}
      
if (p2 !=NULL & n2 != NULL){
     
      n2xp2.assumption.test <- n2*p2
      n2x1minusp2.assumption.test<- n2*(1-p2)

      if (n2xp2.assumption.test<=5 |n2x1minusp2.assumption.test*(1-p2)<=5) {
            warning("WARNING: not valid to assume normal distribution. (n2*p2<5 or n2*(1-p2)<5) ") 
      }
            
      pstar <- (p1+p2)/2
      z <- (p1 - p2)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
      
      #if (alternative=="two.sided"){
      #  2.sided is default    
            p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
            p
      
      #}
      
      if (alternative=="greater"){
       #conditionally evaluated     
            p <- pnorm(z, lower.tail = FALSE)
            p
      
      }
      if (alternative == "less"){
           
            p <- pnorm(z, lower.tail = TRUE)
            p
      }
}
#making confidence interval for p1
#lower <- p1 + qnorm(0) * sqrt(p1 * (1 - p1)/n1)
#upper <- p1 + qnorm(conf.level) * sqrt(p1 * (1 - p1)/n1)
#ci <- c(lower, upper)
#ci
#output<-list(z, p, ci)
#print(output)
}

````
[2] The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity ("MaxLongevity_m") measured in months from species' brain size ("Brain_Size_Species_Mean") measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size).

```{r}
KC<- read.csv("https://raw.githubusercontent.com/difiore/ADA2016/master/KamilarAndCooperData.csv", stringsAsFactors=FALSE, sep=",")
View(KC)
```

### non logged data


Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).
```{r}
library(ggplot2)
hist(KC$MaxLongevity_m)
hist(KC$Brain_Size_Species_Mean)

BL<-lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data=KC)


#so formula is roughly y= 1.22 Bx + 249 
p <- ggplot(data = KC, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m))
p <- p + geom_point()
p <- p + geom_text(aes(x=200,y=750,label=paste("y= 1.22 Bx + 249")))
p <- p + geom_smooth(method = "lm", formula = y ~ x)
p
```

Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

```{r}
SBL<-summary(BL)
```
So, the estimate of slope (B1) is 1.2180. This means that on average,  longevity increases by 1.218 months with every 1 gram increase in brain size. and this significantly different from 0, with a p value of <2e-16. 
```{r}
ci<- confint(BL, level=0.90)
ci
```
There is a 90% probability that the true slope of B1 falls between 1.04 and 1.40


Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.


```{r}
df<-cbind(KC$Brain_Size_Species_Mean, KC$MaxLongevity_m)
Long_Pred<- predict(BL, newdata=data.frame(Brain_Size_Species_Mean=KC$Brain_Size_Species_Mean))# predicts values of longevity given our model and all the brain sizes
Brain_Size<-KC$Brain_Size_Species_Mean
Max_Long<-KC$MaxLongevity_m
df<-cbind(Brain_Size, Max_Long, Long_Pred)
names(df)<- c("b_size", "long", "pred_long")
head(df)
ci <- predict(BL, newdata = data.frame(Brain_Size_Species_Mean=KC$Brain_Size_Species_Mean), interval = "confidence",  level = 0.90)  # conf for all 

pi <- predict(BL, newdata = data.frame(Brain_Size_Species_Mean=KC$Brain_Size_Species_Mean), interval = "prediction",  level = 0.90)  # predict all 
 

df<-cbind(df, ci, pi)
df<-as.data.frame(df)
names(df)<- c("b_size", "long", "pred_long", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr", "PIupr" )
head(df)
 
p <- ggplot(data = df, aes(x = b_size, y =long))
p <- p + geom_point(alpha=1/2)
p <- p + geom_line(aes(x =b_size , y = CIfit), colour = "CIfit")
p <- p + geom_line(aes(x = b_size, y = CIlwr), colour = "CI")
p <- p + geom_line(aes(x = b_size, y = CIupr), colour = "CI")
p <- p + geom_text(aes(x=200,y=750,label=paste("y= 1.22 Bx + 249")))
p <- p + geom_line(data = df, aes(x = b_size, y = PIlwr), colour = "PI")
p <- p + geom_line(data = df, aes(x = b_size, y = PIupr), colour = "PI")
p<- p + scale_colour_manual("", breaks= c("CFIT", "CI" , "CI", "PI", "PI"),   values=c("black", "blue","blue", "yellow" , "yellow"))
p<- p + xlab(" ") +
  scale_y_continuous("Longevity (months)") + 
  labs(title="Relationship between brain size and species mean longevity")
p

p <- ggplot(data = df, aes(x = b_size, y =long))
p <- p + geom_point(alpha=1/2)
p <- p + geom_line(aes(x =b_size , y = CIfit, colour = "predicted values"))+
geom_line(aes(x = b_size, y = CIlwr, colour = "90% confidence interval")) +
 geom_line(aes(x = b_size, y = CIupr, colour = "90% confidence interval"))+
geom_text(aes(x=150,y=790,label=paste("y= 1.22 Bx + 249")))+
geom_line(data = df, aes(x = b_size, y = PIlwr, colour = "90% prediction interval"))+
geom_line(data = df, aes(x = b_size, y = PIupr, colour = "90% prediction interval"))+
 scale_colour_manual("", breaks= c("predicted values", "90% confidence interval" ,  "90% prediction interval"), values=c("pink", "green", "purple"))+
xlab(" ") +
  scale_y_continuous("Longevity (months)") + 
  labs(title="Relationship between brain size and species mean longevity")
p




```



### logged data

hist(log(KC$MaxLongevity_m))
hist(log(KC$Brain_Size_Species_Mean))
